A Wait-Free Queue with Polylogarithmic Step Complexity

### 2023-06-13 presentation/podc-presentation.tex
Some edits to slides

@@

\begin{block}{Our New Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\log^2 p + \log q)$ steps per \op{Dequeue}
\item wait-free
\item uses CAS on reasonable-size words
\item bounded space version: $O(\log p\ \log(p+q))$ amortized steps per operation
(relies on safe GC)
\end{itemize}
\end{block}

@@

% 60% size; others are 80%
\hspace*{7mm}
\only<1>{\input{OrderTree-5}}% 50-53
\only<2>{\input{OrderTree-6}}% 48,50-53
\only<3>{\input{OrderTree-7}}% 50-54
\only<4>{\input{OrderTree-8}}% 50-55
\only<5>{\input{OrderTree-9}}% 50-56

@@

\begin{block}{New Wait-Free Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\mbox{log}^2 p + \log q)$ steps per \op{Dequeue}
\item<2> Unbounded space
\end{itemize}
\end{block}

@@

\begin{itemize}
\item Replace array of blocks with a red-black tree of blocks
\item Periodically split RBT and discard obsolete blocks
\item Processes  help one another to ensure blocks are obsolete
\end{itemize}

>>>>

\begin{frame}{Queues:  Breaking Linear-Time Bottleneck}

Problem:  implement lock-free FIFO queue
\begin{itemize}
\item shared by $p$ processes
\item use single-word CAS (reasonable-sized words)
\item support multiple enqueuers, dequeuers
\end{itemize}

\medskip
\pause

Many previous solutions for this problem.

All require $\Omega(p)$ steps per operation\\
$\rightarrow$ Real obstacle to scalability

\pause
\begin{block}{Our New Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\mbox{log}^2 p + \log q)$ steps per \op{Dequeue} ($q=$ size of queue)
\item wait-free
\end{itemize}
\end{block}

@@

\begin{block}{Our New Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\mbox{log}^2 p + \log q)$ steps per \op{Dequeue}
\item wait-free
\item uses CAS on reasonable-size words
\item bounded space version: $O(\log p\ \log(p+q))$ amortized steps per operation
(relies on safe GC)
\end{itemize}
\end{block}

@@

% 60% size; others are 80%
%\hspace*{7mm}
\only<1>{\input{OrderTree-5}}% 50-53
\only<2>{\input{OrderTree-6}}% 48,50-53
\only<3>{\input{OrderTree-7}}% 50-54
\only<4>{\input{OrderTree-8}}% 50-55
\only<5>{\input{OrderTree-9}}% 50-56

@@

\begin{block}{New Wait-Free Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\mbox{log}^2 p + \log q)$ steps per \op{Dequeue}
\item $O(\log p)$ CAS steps per \op{Dequeue}
\item<2> Unbounded space
\end{itemize}
\end{block}

@@

\begin{itemize}
\item Replace each array of blocks with a red-black tree of blocks
\item Periodically split RBT and discard obsolete blocks
\item Processes  help one another to ensure blocks are obsolete
\end{itemize}
